---
title: "SMAGL Notebook - Spring 2021"
author: "Nick Mikulski"
date: "2/5/2021"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: yeti
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction
This document will serve as my lab notebook for SMAGL research under Prof. Christopher Schmitt during the Spring 2021 semester. My research for credit application was approved on Feb. 5, 2021.
  
[Application](https://docs.google.com/document/d/1ODJpSP9Owl21iwok2X3M-f-7nAIKtkkMasPKDSp8i4g/edit?usp=sharing)


## Goals for the semester
Given that my main academic interests are in conservation biology, I look forward to being able to work with Laura Angley and assist with her research on the release of rehabilitated vervet monkey populations.  
  
- Learn how to conduct a social network analysis in R  
- Using the rehabilitated vervet monkey dataset, analyze the social networks of the pre-release and post-release populations
- Compare pre-release and post-release networks to look for patterns such as rank and centrality and to identify possible vulnerabilities in the network

# Learning to use SNA R Packages

## Walk through of: McFarland *et al.* 2010    
[McFarland DA, Messing S, Nowak M, Westwood SJ. (2010) Network Analysis Labs in R and SoNIA.](https://sna.stanford.edu/rlabs.php)  
    

**Inputting Data from website**   
source("http://sna.stanford.edu/setup.R")    
install.packages("igraph")   
library(igraph)   

```{r}
advice_data_frame <- read.table('http://sna.stanford.edu/sna_R_labs/data/Krack-High-Tec-edgelist-Advice.txt')
friendship_data_frame <- read.table('http://sna.stanford.edu/sna_R_labs/data/Krack-High-Tec-edgelist-Friendship.txt')
reports_to_data_frame <- read.table('http://sna.stanford.edu/sna_R_labs/data/Krack-High-Tec-edgelist-ReportsTo.txt')

head(advice_data_frame)

attributes <- read.csv('http://sna.stanford.edu/sna_R_labs/data/Krack-High-Tec-Attributes.csv', header=T)
attributes
```

Adding column names for organization
```{r}
colnames(advice_data_frame) <- c('ego', 'alter', 'advice_tie')
head(advice_data_frame)
  
colnames(friendship_data_frame) <- c('ego', 'alter', 'friendship_tie')
head(friendship_data_frame)
  
colnames(reports_to_data_frame) <- c('ego', 'alter', 'reports_to_tie')
head(reports_to_data_frame)
```

These lines open the partial data frames in an XQuartz window, but I took out of chunk because it was messing up the knitting of the file.   
    
fix(advice_data_frame)   
fix(friendship_data_frame)   
fix(reports_to_data_frame)    

Confirms that our columns and rows align before combining the data set:
```{r}
advice_data_frame$ego == friendship_data_frame$ego

which(advice_data_frame$ego != friendship_data_frame$ego)
which(advice_data_frame$alter != friendship_data_frame$alter)
which(reports_to_data_frame$alter != friendship_data_frame$alter)
which(reports_to_data_frame$ego != friendship_data_frame$ego)
```

Combining them into one data frame:
```{r}
krack_full_data_frame <- cbind(advice_data_frame, 
    friendship_data_frame$friendship_tie, 
    reports_to_data_frame$reports_to_tie)
head(krack_full_data_frame)
```

Rename columns to organize better:
```{r}
names(krack_full_data_frame)[4:5] <- c("friendship_tie", 
    "reports_to_tie")  
head(krack_full_data_frame)

krack_full_nonzero_edges <- subset(krack_full_data_frame, 
    (advice_tie > 0 | friendship_tie > 0 | reports_to_tie > 0))
head(krack_full_nonzero_edges)
```

Reduce to only non-zero edges to only contain actual connections:
```{r}
library(igraph) #Unsure why, but it fails every time if not reloaded within the chunk
krack_full <- graph.data.frame(krack_full_nonzero_edges) 
summary(krack_full)
krack_full
```

Create vectors of specific edge types:
```{r}
get.edge.attribute(krack_full, 'advice_tie')
get.edge.attribute(krack_full, 'friendship_tie')
get.edge.attribute(krack_full, 'reports_to_tie')
```
Symmetrize the network, making all asymmetric ties symmetric:
```{r}
krack_full_symmetrized <- as.undirected(krack_full, mode='collapse')
summary(krack_full_symmetrized)
```

**Adding Attributes to the Graph**    
Create a vector of vertex labels
```{r}
attributes = cbind(1:length(attributes[,1]), attributes) #I'm a little confused on the arguments within esp the [,1]
krack_full <- graph.data.frame(d = krack_full_nonzero_edges, 
                               vertices = attributes)


#Attributes (age, tensure, level, and dept) are now listed alongside name as vertex attributes
summary(krack_full)
```
?cbind

List all values for a given attribute
```{r}
get.vertex.attribute(krack_full, 'AGE')
get.vertex.attribute(krack_full, 'TENURE')
get.vertex.attribute(krack_full, 'LEVEL')
get.vertex.attribute(krack_full, 'DEPT')
```

    
**Visualizing the network**   
Plot of initial network
```{r}
plot(krack_full)

```
    
Saves plotted graph as a pdf:    
```{r}
setwd("/users/nickmikulski/Desktop/SMAGL/Spring2021") #set working directory to define where pdf will be save to
pdf("1.1_Krackhardt_Full.pdf")
plot(krack_full)
dev.off() #this line ends the pdf save process?

```

?delete.edges
  (graph, edges-to-be-deleted)

?E

?get.edge.attribute

This network is very messy and unclear, so the lines below reduce to just one edge type in the network.   
```{r}
#First, just the advice edges
krack_advice_only <- delete.edges(krack_full, #delete edges within krack_full:
    E(krack_full)[get.edge.attribute(krack_full, #E(krack_full) gives all edges; [get. argument keeps just the advice edges
    name = "advice_tie") == 0])
summary(krack_advice_only)
plot(krack_advice_only) #still very messy

#Friendship edges only
krack_friendship_only <- delete.edges(krack_full, 
    E(krack_full)[get.edge.attribute(krack_full, 
    name = "friendship_tie") == 0])
summary(krack_friendship_only)
plot(krack_friendship_only)

#Reports to edges only
krack_reports_to_only <- delete.edges(krack_full, 
    E(krack_full)[get.edge.attribute(krack_full, 
    name = "reports_to_tie") == 0])
summary(krack_reports_to_only)
plot(krack_reports_to_only) #This is very interesting. It is definitely the most orderly network, which makes sense for "reports to."
```

I hope to be able to create a network similar to this with our real data to be able to see which individual vervets are central to the group.    

In order to further visually clean up the networks, we can use a preset layout on our advice-only network
I will be following the layout Fruchterman-Rheingold just to follow along with this source, but others can be found at ?layout.    
```{r}
reports_to_layout <- layout.fruchterman.reingold(krack_reports_to_only)
plot(krack_reports_to_only, 
     layout=reports_to_layout) #All this really did was re-order the primary connections with the largest sub-networks at bottom
```

Further customize plot    
First, make a vector of colors.   
```{r}
dept_vertex_colors = get.vertex.attribute(krack_full,"DEPT") #color-coding by department
colors = c('Black', 'Red', 'Blue', 'Yellow', 'Green')
dept_vertex_colors[dept_vertex_colors == 0] = colors[1]
dept_vertex_colors[dept_vertex_colors == 1] = colors[2]
dept_vertex_colors[dept_vertex_colors == 2] = colors[3]
dept_vertex_colors[dept_vertex_colors == 3] = colors[4] 
dept_vertex_colors[dept_vertex_colors == 4] = colors[5]

plot(krack_reports_to_only, 
    layout=reports_to_layout, 
    vertex.color=dept_vertex_colors, #color-coding vertices by vector above
    vertex.label=NA, #removing vertex labels
    edge.arrow.size=.5) #making the arrows smaller

tenure_vertex_sizes = get.vertex.attribute(krack_full,"TENURE") #changing vertex size to represent tenure attribute
 
plot(krack_reports_to_only, 
     layout=reports_to_layout, 
     vertex.color=dept_vertex_colors, 
     vertex.label=NA, 
     edge.arrow.size=.5,
     vertex.size=tenure_vertex_sizes)
```
      
That was really satisfying to be able to customize the visuals of the network.   
Next, I will overlay this network with the other edge types: advice & friendships in red and blue using the same layout   

?rgb (red, blue, alpha, names=NULL, maxColorValue=1) where alpha is opacity (0=transparent, 1=opaque)

```{r}
tie_type_colors = c(rgb(1,0,0,.5), rgb(0,0,1,.5), rgb(0,0,0,.5)) #creates red, blue, grey(?)
E(krack_full)$color[ E(krack_full)$advice_tie==1 ] = tie_type_colors[1]
E(krack_full)$color[ E(krack_full)$friendship_tie==1 ] = tie_type_colors[2]
E(krack_full)$color[ E(krack_full)$reports_to_tie==1 ] = tie_type_colors[3]
E(krack_full)$arrow.size=.5 
V(krack_full)$color = dept_vertex_colors
V(krack_full)$frame = dept_vertex_colors
 
plot(krack_full, 
     layout=reports_to_layout, 
     vertex.color=dept_vertex_colors, 
     vertex.label=NA, 
     edge.arrow.size=.5,
     vertex.size=tenure_vertex_sizes)

```
   
Lastly, I will add a legend to improve read-ability (place legend function immediately after)
```{r}
plot(krack_full, 
     layout=reports_to_layout, 
     vertex.color=dept_vertex_colors, 
     vertex.label=NA, 
     edge.arrow.size=.5,
     vertex.size=tenure_vertex_sizes)
legend(1, 
       1.25,
       legend = c('Advice', 
                  'Friendship',
                  'Reports To'), 
       col = tie_type_colors, 
       lty=1,
       cex = .7)

```

Yay! This network is complete. I found this walk-through to be really valuable in preparing to begin constructing a social network of our vervet monkey dataset. I think this will help me a lot with 


## Walk through of/Notes on: Sadler 2017    
[Sadler J. (2017) Introduction to Network Analysis with R.](https://www.jessesadler.com/post/network-analysis-with-r/)      




# Beginning with Vervet Dataset    
## Getting started with help from Laura   
[Pre-Release Social Proximity Dataset](https://raw.githubusercontent.com/nickmikulski/Spring2021/main/Pre-release_Social%20Proximity_csv.csv)    

Loading necessary packages (File would not knit without putting these in a chunk)    
```{r}
library(curl)
library(dplyr)
library(tidyverse)

```

### Inputting Datasets
```{r}
pre_socnet<- curl("https://raw.githubusercontent.com/nickmikulski/Spring2021/main/Pre-release_Social%20Proximity_csv.csv")
pre_socnet<- read.csv(pre_socnet, header = T, na.strings=c(""," ","NA"))
#head(pre_socnet) #check the data to make sure it's uploaded correctly
#str(pre_socnet)

post_socnet<- curl("https://raw.githubusercontent.com/langley1/LWTdata2016/main/Post-release_Social%20Proximity.csv")
post_socnet<- read.csv(post_socnet, header = T, na.strings=c(""," ","NA"))
#head(post_socnet) #check the data to make sure it's uploaded correctly

```
   
### Editting the dataframe
```{r}
pre_socnet_close<- pre_socnet %>% #creating a new dataframe called pre_socnet_close using data from the original pre_socnet dataframe
  filter(Focal.ID != "BT", #this code REMOVES all data that has Batman has the focal ID (BT is wild male from prerelease)
         Association != "BT", #this code REMOVES all data that has Batman in association column
         Proximity.Code %in% c("1","2") #this code only keeps proximity codes 1,2 (excluding 3,4) because we are focusing on closer proximity
  ) 


```

### Taking out a sub-sample from dataframe
This section pulls out a few specific individuals in order to practice coding using the dataframe
```{r}
pre_socnet_sample<- pre_socnet_close %>%
  filter(Focal.ID %in% c("PO","BL","KO","ZI","AM"), Association %in% c("PO","BL","KO","ZI","AM"))

```

### Creating a matrix of sub-sample
```{r}
# 1. Create a character vector of all the focal IDs in dataset:
pre_sn_sample_IDs<-as.character(unique(pre_socnet_sample$Focal.ID))
pre_sn_sample_IDs
```

```{r}
# 2. Get a list of dataframes, subsetted by monkey ID:
pre_sn_sample_monkeylist<-lapply(pre_sn_sample_IDs, function(x){pre_socnet_sample[pre_socnet_sample[["Focal.ID"]] == x, ]})
# The line above is a little bit confusing. It is creating a separate dataframe for each individual based on their focal id
#head(pre_sn_sample_monkeylist) #hidden for knitting
```

```{r}
# 3. Group each by focal/associate, and count how many times they are observed close together:
pre_sn_sample_grouped<-
  pre_sn_sample_monkeylist %>%
  purrr::map(~group_by(.,Association)) %>%
  purrr::map(~summarize(.,count=n())) 
#pre_sn_sample_grouped

```

?purrr::map # This is pulling the map command from the purrr package    
?~ # The tilde (dependentvariable~independent)   

```{r}
names(pre_sn_sample_grouped) <- pre_sn_sample_IDs #this gives each grouped list the name of the Focal ID
#pre_sn_sample_grouped

# 4. Set up pairwise combinations of interacting monkeys:
pre_sn_sample_monkeycombos<-list(focal=pre_sn_sample_IDs, associate=pre_sn_sample_IDs) #create list of all possible focals/associates
pre_sn_sample_filt<- function(x, y) {x == y} #create function to filter out same-monkey pairs ("PO is close to PO")
pre_sn_sample_combo<- pre_sn_sample_monkeycombos %>% cross_df(.,.filter=pre_sn_sample_filt) #get the filtered combined list as a dataframe
#head(pre_sn_sample_combo)
```

?cross_df #

```{r}
# 5. Create new dataframes with specific criteria
pre_sn_sample_combo2<-
  pre_sn_sample_combo %>%
  mutate(absent1 = map2_chr( #new column called "absent1"
    focal,
    associate,
    ~if_else(.x %in% names(pre_sn_sample_grouped),true="TRUE",false="FALSE"))) %>%
    mutate(absent2 = map2_chr(
    focal,
    associate,
    ~if_else(.y %in% pre_sn_sample_grouped[[.x]]$Association,true="TRUE",false="FALSE"))) %>%
  filter(absent1 == "TRUE") %>%
  filter(absent2 == "TRUE") %>%
  dplyr::select(-absent1,-absent2) #this removes those two new columns you made so you're just left with the ID names

pre_sn_sample_combo3<- pre_sn_sample_combo2 %>% 
  mutate(proximity = map2_int( #new column called "proximity" that is the count for when proximity code = 1 or 2
    focal, 
    associate, 
    ~pre_sn_sample_grouped %>% pluck(.x) %>% filter(Association==.y) %>% as.data.frame(.) %>% .[,2]))
head(pre_sn_sample_combo3)

```
Finally, create the matrix    
```{r}
# 6. Create your matrix
pre_sn_sample_matrix<-spread(pre_sn_sample_combo3,associate,proximity) %>% column_to_rownames(var="focal") %>% data.matrix()
pre_sn_sample_matrix

```

?spread

## Creating a social network plot based on this matrix  
### Converting the matrix into a data frame     

?as.data.frame



```{r}
pre_sn_sample_df <- as.data.frame(pre_sn_sample_matrix, row.names = pre_sn_sample_IDs, colnames(pre_sn_sample_IDs), stringsAsFactors = TRUE)
pre_sn_sample_df
```

Plotting this newly formed data frame:
```{r}
library(igraph)

pre_sn_sample_edges <- pre_sn_sample_df
pre_sn_sample_nodes <- pre_sn_sample_IDs
```

```{r}
get.vertex.attribute(pre_sn_sample_df, 'DEPT')
```

?graph_from_data_frame

?graph.data.frame

```{r}
library(igraph)
pre_sn_sample_graph <- pre_sn_sample_graph <- graph_from_data_frame(d=pre_sn_sample_edges, vertices = pre_sn_sample_nodes, directed = TRUE) #Using the matrix

pre_sn_sample_graph

```
Alternatives (?):    
pre_sn_sample_graph <- graph_from_data_frame(d=pre_sn_sample_edges, vertices = pre_sn_sample_nodes, directed = TRUE) #Using the matrix

pre_sn_sample_graph <- graph_from_data_frame(d=(as.data.frame(pre_sn_sample_combo3)), vertices = pre_sn_sample_nodes, directed = TRUE) #using the step before the matrix


?plot

```{r}
plot(pre_sn_sample_graph)


```












# Citations

## R Coding References

- [Barbera P. (2017) Introduction to social network analysis with R.](http://pablobarbera.com/big-data-upf/html/02a-networks-intro-visualization.html)  
- [McFarland DA, Messing S, Nowak M, Westwood SJ. (2010) Network Analysis Labs in R and SoNIA.](https://sna.stanford.edu/rlabs.php)  
- [Sadler J. (2017) Introduction to Network Analysis with R.](https://www.jessesadler.com/post/network-analysis-with-r/)  
- [Zhang G. (2010) Social network analysis with R sna package.](https://www.r-project.org/nosvn/conferences/useR-2010/slides/Zhang.pdf)  


## Social Network Analysis Literature

- [Brent LJN, Lehmann J, Ramos-Fernández G. (2011) Social network analysis in the study of nonhuman primates: A historical perspective. *American Journal of Primatology*, **73**, 720-730.](https://doi.org/10.1002/ajp.20949)  
- [Guy AJ, Stone OML, Curnoe D. (2012) The release of a troop of rehabilitated vervet monkeys (Chlorocebus aethiops) in KwaZulu-Natal, South Africa: Outcomes and Assessment. *Folia Primatologica*, **82**, 308-320.](https://doi.org/10.1159/000337269)  
- [McCowan B, Anderson K, Heagarty A, Cameron A. (2008) Utility of social network analysis for primate behavioral management and well-being. *Applied Animal Behaviour Science*, **109**, 396-405.](https://doi.org/10.1016/j.applanim.2007.02.009)  
- [Puga-Gonzalez I, Sosa S, Sueur C. (2019) Editorial: Social networks analyses in primates, a multilevel perspective. *Primates*, **60**, 163-165.](https://doi.org/10.1007/s10329-019-00720-5)  
- [Sueur C, Jacobs A, Amblard F, Petit O, King AJ. (2011) How can social network analysis improve the study of primate behavior? *American Journal of Primatology*, **73**, 703-719.](https://doi.org/10.1002/ajp.20915)  